The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2023-09-27 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[twocolumn,linenumbers,twocolappendix]{aastex631}

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{layouts}
\usepackage{xcolor}
\usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}

@@

% Author list
\author[0000-0003-3781-0747]{Liam O. Dubay}
\author[0000-0001-7258-1834]{Jennifer A. Johnson}
\author[0000-0002-6534-8783]{James W. Johnson}
% \author[0000-0001-7775-7261]{David H. Weinberg}

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters which have been left unchanged from \citet{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & & Maximum simulation time \\
        n               & 8         & & Number of stellar populations formed per ring per time-step \\
        R_{\rm Ia}(t)   & Equations \ref{eq:powerlaw-dtd}--\ref{eq:triple-dtd}  & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        M_{\rm gas,0}   & 0         & & Initial gas mass \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & & SN Ia yield of Fe \\
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & & Time-dependence of the star formation rate in the inside-out model \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & & Time-dependence of the star formation rate in the late-burst model \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & & Time-dependence of the infall rate in the early-burst model \\
        \uptau_{*,\rm EB}(t|R_{\rm gal})  & Equation \ref{eq:earlyburst-conroy22} & & Time-dependence of the SFE timescale in the early-burst model \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & & Time-dependence of the infall rate in the two-infall model \\
    \enddata
\end{deluxetable*}

@@

We explore five different functional forms for the SN Ia DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, a two-population model, and a model based on triple-system dynamics, with a one or two different parameterizations for each. Figure \ref{fig:dtds} presents these DTDs as a function of time after star formation.

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Models of the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized to $f_{\rm Ia}=1$ at $\tau=1$ Gyr. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD}, to the same scale as the DTD models. The horizontal error bars indicate the time range of each DTD measurement, and the vertical error bars represent the 1$\sigma$ uncertainties.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    R(t) = A_\alpha (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is a standard assumption in many observational studies of the DTD as well as GCE simulations. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Plateau} A ``plateau'' model consisting of a flat slope of width $W$ followed by a power law decline with normalization $B_W$:
\begin{equation}
    R(t) = 
    \begin{cases}
        B_W, & t < W \\
        B_W (t/W)^\alpha, & t \ge W
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models. Figure \ref{fig:analytical-dtd} illustrates the similarity between these models and two different treatments of the analytical double-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. For more discussion on this, see Appendix \ref{app:analytical-dtds}.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$ and normalization $C_\tau$:
\begin{equation}
    R(t) = \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr. \citet{Matteucci1986-SupernovaEnrichment}, \citet{Stolger2020-ExponentialDTD}

\paragraph{Two-population} A DTD in which 50\% of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the other 50\% form an exponential tail at large $t$, normalized by $D_{t_p,\sigma,\tau}$:
\begin{equation}
    R(t) = D_{t_p,\sigma,\tau} \Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\tau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
Following the two-population model from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\tau=3$ Gyr, which results in roughly 50\% of SNe Ia exploding within $t<100$ Myr. \citet{Poulhazan2018-PrecisionPollution} refer to this as the ``bimodal'' model.

\paragraph{Triple} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2022-TripleEvolution}. We approximate their DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    R(t) = 
    \begin{cases}
        \epsilon B_W, & t < t_{\rm rise} \\
        B_W, & t_{\rm rise} \leq t < W \\
        B_W (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
We take $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

Table \ref{tab:dtds} summarizes the model DTDs we investigate in this paper.

@@

\begin{figure*}
    \centering
    \includegraphics{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_*$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_{\rm gas}$ (third row), and the star formation efficiency timescale $\tau_*$ (fourth row) as functions of simulation time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the simulation zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/feh_df_comparison.pdf}
    \caption{Distributions of [Fe/H] from multi-zone simulations with various models for the SFH and DTD. Each row presents distributions of stars within a range of absolute midplane distance: $1\leq|z|<2$ kpc (\textit{top}), $0.5\leq|z|<1$ kpc (\textit{middle}), and $0\leq|z|<0.5$ kpc (\textit{bottom}). Within each panel, curves of different color represent the distributions of stars binned by Galactocentric radius $R_{\rm gal}$, from $3\leq R_{\rm gal}<5$ kpc (yellow) to $13\leq R_{\rm gal}<15$ kpc (blue). Each distribution is normalized so the area under the curve is 1, and the vertical scale is consistent across each row. All distributions are convolved with observational uncertainties in APOGEE DR17 (see Table \ref{tab:uncertainties}) and smoothed with a box-car width of 0.2 dex. 
    \textit{Left columns:} comparison between the inside-out and two-infall SFH models. In both cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. 
    \textit{Center column:} the distributions from APOGEE DR17 for reference, binned and smoothed similarly.
    \textit{Right columns:} comparison between the power-law ($\alpha=-1.4$) and exponential ($\tau=3$ Gyr) DTD models. In both cases the inside-out SFH is assumed.}
    \label{fig:feh-df-comparison}
    \script{feh_df_comparison.py}
\end{figure*}

@@

% \begin{figure*}
%     \centering
%     \includegraphics{figures/feh_df_sfh.pdf}
%     \caption{Distributions of [Fe/H] from multi-zone simulations with different SFHs. The plot format is similar to Figure \ref{fig:feh-df-dtd}. In all cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. Distributions from APOGEE DR17, binned and smoothed similarly, are presented in the right-most column for reference.}
%     \label{fig:feh-df-sfh}
%     \script{feh_df_sfh.py}
% \end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/ofe_df_sfh.pdf}
    \caption{Distributions of [O/Fe] from multi-zone simulations with different SFHs. In all cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. The format of each panel is the same as in Figure \ref{fig:feh-df-comparison}, except that all distributions are smoothed with a box-car width of 0.05 dex. Distributions from APOGEE DR17, binned and smoothed similarly, are presented in the right-most column for reference.}
    \label{fig:ofe-df-sfh}
    \script{ofe_df_sfh.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/ofe_df_dtd.pdf}
    \caption{Distributions of [O/Fe] from multi-zone simulations with different DTDs. In all cases an inside-out SFH is assumed. The plot format is similar to Figure \ref{fig:ofe-df-sfh}.}
    \label{fig:ofe-df-dtd}
    \script{ofe_df_dtd.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/ofe_bimodality_summary.pdf}
    \caption{The distributions of [O/Fe] along two slices of [Fe/H]: $-0.6\leq$[Fe/H]$<-0.4$ (blue dashed) and $-0.4\leq$[Fe/H]$<-0.2$ (red solid). \textit{Top row:} results from five multi-zone simulations which assume the late-bust SFH but different DTD models. \textit{Bottom row}: the first four panels compare the four SFH models, all assuming an exponential DTD with $\tau=1.5$ Gyr. The bottom-right panel plots data from APOGEE DR17 for reference. All panels contain stars within the Galactic region defined by $7\leq R_{\rm gal}<9$ kpc and $0\leq|z|<2$ kpc. The distributions of stars in $|z|$ The maximum of each distribution is normalized to 1 and the vertical scale is consistent across all panels.}
    \label{fig:ofe-bimodality}
    \script{ofe_bimodality_summary.py}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_sfh.pdf}
    \caption{A comparison of the [O/Fe]--[Fe/H] plane between the four SFH models in our multi-zone simulations. All assume an exponential DTD with $\tau=1.5$ Gyr. Each panel plots a random mass-weighted sample of 10,000 star particles in the Solar neighborhood ($7\leq R_{\rm gal}<9$ kpc, $0\leq|z|<0.5$ kpc) color-coded by the Galactocentric radius at birth. A Gaussian scatter has been applied to all points based on the median abundance errors in APOGEE DR17 (ses Table \ref{tab:uncertainties}). The black curves represent the ISM abundance tracks in the 8 kpc zone. The red contours represent the APOGEE abundance distribution, with the solid and dashed contours enclosing 68\% and 95\% of stars in the sample, respectively.}
    \label{fig:ofe-feh-sfh}
    \script{ofe_feh_sfh.py}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics{figures/ofe_feh_dtd.pdf}
    \caption{The [O/Fe]--[Fe/H] plane from multi-zone simulations with different DTD models. All assume the inside-out SFH. Each panel is similar to those in Figure \ref{fig:ofe-feh-sfh}, except each row contains star particles from a different bin in $|z|$, with stars closest to the midplane in the bottom row and stars farthest from the midplane in the top row. All panels present stars within the Solar annulus ($7\leq R_{\rm gal}<9$ kpc).}
    \label{fig:ofe-feh-dtd}
    \script{ofe_feh_dtd.py}
\end{figure*}

@@

@@

\begin{acknowledgments}

Funding for the Sloan Digital Sky 
Survey IV has been provided by the 
Alfred P. Sloan Foundation, the U.S. 
Department of Energy Office of 
Science, and the Participating 
Institutions.

>>>>

% Define document class
\documentclass[modern,linenumbers]{aastex631}
%twocolappendix

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
\usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}

\shorttitle{Galactic SN Ia DTD}
\shortauthors{Dubay, Johnson, \& Johnson}

@@

% Author list
\author[0000-0003-3781-0747]{Liam O. Dubay}
\affiliation{\osuaffil}
\author[0000-0001-7258-1834]{Jennifer A. Johnson}
\affiliation{\osuaffil}
\affiliation{Center for Cosmology and Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\author[0000-0002-6534-8783]{James W. Johnson}
\affiliation{\osuaffil}
\affiliation{Carnegie Observatories}
% \author[0000-0001-7775-7261]{David H. Weinberg}

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters which have been left unchanged from \citet{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & & Maximum simulation time \\
        n               & 8         & & Number of stellar populations formed per ring per time-step \\
        R_{\rm Ia}(t)   & Equations \ref{eq:powerlaw-dtd}--\ref{eq:triple-dtd}  & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        M_{\rm gas,0}   & 0         & & Initial gas mass \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & & SN Ia yield of Fe \\
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & & Time-dependence of the SFR in the inside-out model \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & & Time-dependence of the SFR in the late-burst model \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & & Time-dependence of the IFR in the early-burst model \\
        \uptau_{*,\rm EB}(t|R_{\rm gal})  & Equation \ref{eq:earlyburst-conroy22} & & Time-dependence of the SFE timescale in the early-burst model \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & & Time-dependence of the IFR in the two-infall model \\
    \enddata
\end{deluxetable*}

@@

\begin{subequations}
\begin{align}
    \dot M_{\rm Fe}^{\rm Ia} &= y_{\rm Fe}^{\rm Ia} \langle \dot M_*\rangle_{\rm Ia} \\
    &= y_{\rm Fe}^{\rm Ia} \frac{\int_0^t \dot M_*(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
\end{align}
\end{subequations}

The yield $y_{\rm Fe}^{\rm Ia}$ can be expressed in terms of the integral of the DTD:

\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_*}
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia. We take the average number of SNe Ia per mass of stars formed to be $N_{\rm Ia}/M_*=2.2\times10^{-3}\,{\rm M}_\odot^{-1}$ following \citet{MaozMannucci2012-SNeIaReview}, which gives $R_{\rm Ia}$ units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$.

We explore five different functional forms for the SN Ia DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, a two-population model, and a model based on triple-system dynamics, with a one or two variants of each. We present the unnormalized functional forms for $R_{\rm Ia}(t)$ below; VICE automatically normalizes and scales the DTD to produce the correct number of SNe Ia over the lifetime of the Galaxy. Figure \ref{fig:dtds} presents these DTDs as a function of time after star formation, and Table \ref{tab:dtds} summarizes the DTD models and their parameters.

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $R_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTD models. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    R_{\rm Ia}(t) \propto (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
% where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is a standard assumption in many observational studies of the DTD as well as GCE simulations. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Plateau} A ``plateau'' model consisting of a flat slope of width $W$ followed by a power law decline:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
% where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models. Figure \ref{fig:analytical-dtd} illustrates the similarity between these models and two different treatments of the analytical double-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. For more discussion on this, see Appendix \ref{app:analytical-dtds}.

\paragraph{Triple-system} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2022-TripleEvolution}. We approximate their DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        \epsilon, & t < t_{\rm rise} \\
        1, & t_{\rm rise} \leq t < W \\
        (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
We take $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    R_{\rm Ia}(t) \propto \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous GCE studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr. \citet{Matteucci1986-SupernovaEnrichment}, \citet{Stolger2020-ExponentialDTD}

\paragraph{Two-population} A DTD in which 50\% of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the other 50\% form an exponential tail at large $t$:
\begin{equation}
    R_{\rm Ia}(t) \propto 0.5\Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\tau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
Following the two-population model from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\tau=3$ Gyr, which results in roughly 50\% of SNe Ia exploding within $t<100$ Myr. \citet{Poulhazan2018-PrecisionPollution} refer to this as the ``bimodal'' model.

% Minimum SN Ia delay time
The minimum SN Ia delay time $t_D$, i.e. the lifetime of the most massive progenitor system, is also a parameter of the DTD which can have an effect on chemical evolution. \citet{Andrews2017-ChemicalEvolution} found that a longer minimum delay time produces a sharper high-$\alpha$ ``knee'' and shifts it to higher [Fe/H], although they only explored this for an exponential DTD. They adopt a fiducial value of $t_D=150$ Myr, as do \citet{Johnson2021-Migration}. \citet{Poulhazan2018-PrecisionPollution} adopt a much smaller value of $t_D=29$ Myr as the lifetime of their most massive progenitors.

We take $t_D=40$ Myr to be our fiducial value as it is the approximate lifetime of an $8\,{\rm M}_\odot$ star [Larson 1974]. We find that adopting a longer $t_D$ has only a minor effect on the chemical evolution for most DTD models except the power-law, but in that case the effect of a longer minimum delay time can be approximated by adding an initial plateau of $\sim100$ Myr to the DTD. See Section \ref{sec:onezone-results} and Figure \ref{fig:onezone-twopanel} for details.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_*$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_{\rm gas}$ (third row), and the star formation efficiency timescale $\tau_*$ (fourth row) as functions of simulation time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the simulation zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/feh_df_comparison.pdf}
    \caption{Distributions of [Fe/H] from multi-zone simulations with various models for the SFH and DTD. Each row presents distributions of stars within a range of absolute midplane distance: $1\leq|z|<2$ kpc (\textit{top}), $0.5\leq|z|<1$ kpc (\textit{middle}), and $0\leq|z|<0.5$ kpc (\textit{bottom}). Within each panel, curves of different color represent the distributions of stars binned by Galactocentric radius $R_{\rm gal}$, from $3\leq R_{\rm gal}<5$ kpc (yellow) to $13\leq R_{\rm gal}<15$ kpc (blue). Each distribution is normalized so the area under the curve is 1, and the vertical scale is consistent across each row. All distributions are convolved with observational uncertainties in APOGEE DR17 (see Table \ref{tab:uncertainties}) and smoothed with a box-car width of 0.2 dex. 
    \textit{Left columns:} comparison between the inside-out and two-infall SFH models. In both cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. 
    \textit{Center column:} the distributions from APOGEE DR17 for reference, binned and smoothed similarly.
    \textit{Right columns:} comparison between the power-law ($\alpha=-1.4$) and exponential ($\tau=3$ Gyr) DTD models. In both cases the inside-out SFH is assumed.}
    \label{fig:feh-df-comparison}
    \script{feh_df_comparison.py}
\end{figure*}

@@

% \begin{figure*}
%     \centering
%     \includegraphics[width=\linewidth]{figures/feh_df_sfh.pdf}
%     \caption{Distributions of [Fe/H] from multi-zone simulations with different SFHs. The plot format is similar to Figure \ref{fig:feh-df-dtd}. In all cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. Distributions from APOGEE DR17, binned and smoothed similarly, are presented in the right-most column for reference.}
%     \label{fig:feh-df-sfh}
%     \script{feh_df_sfh.py}
% \end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/ofe_df_sfh.pdf}
    \caption{Distributions of [O/Fe] from multi-zone simulations with different SFHs. In all cases an exponential DTD with timescale $\tau=1.5$ Gyr is assumed. The format of each panel is the same as in Figure \ref{fig:feh-df-comparison}, except that all distributions are smoothed with a box-car width of 0.05 dex. Distributions from APOGEE DR17, binned and smoothed similarly, are presented in the right-most column for reference.}
    \label{fig:ofe-df-sfh}
    \script{ofe_df_sfh.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/ofe_df_dtd.pdf}
    \caption{Distributions of [O/Fe] from multi-zone simulations with different DTDs. In all cases an inside-out SFH is assumed. The plot format is similar to Figure \ref{fig:ofe-df-sfh}.}
    \label{fig:ofe-df-dtd}
    \script{ofe_df_dtd.py}
\end{figure*}

@@

Reference \citet{Vincenzo2021-AlphaDistribution} for how the bimodality is made stronger by the selection function, but it's not just due to selection effects.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/ofe_bimodality_summary.pdf}
    \caption{The distributions of [O/Fe] along two slices of [Fe/H]: $-0.6\leq$[Fe/H]$<-0.4$ (blue dashed) and $-0.4\leq$[Fe/H]$<-0.2$ (red solid). \textit{Top row:} results from five multi-zone simulations which assume the late-bust SFH but different DTD models. \textit{Bottom row}: the first four panels compare the four SFH models, all assuming an exponential DTD with $\tau=1.5$ Gyr. The bottom-right panel (highlighted) plots data from APOGEE DR17 for reference. All panels contain stars within the Galactic region defined by $7\leq R_{\rm gal}<9$ kpc and $0\leq|z|<2$ kpc. The distributions of stars in $|z|$ The maximum of each distribution is normalized to 1 and the vertical scale is consistent across all panels.}
    \label{fig:ofe-bimodality}
    \script{ofe_bimodality_summary.py}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_sfh.pdf}
    \caption{A comparison of the [O/Fe]--[Fe/H] plane between the four SFH models in our multi-zone simulations. All assume an exponential DTD with $\tau=1.5$ Gyr. Each panel plots a random mass-weighted sample of 10,000 star particles in the Solar neighborhood ($7\leq R_{\rm gal}<9$ kpc, $0\leq|z|<0.5$ kpc) color-coded by the Galactocentric radius at birth. A Gaussian scatter has been applied to all points based on the median abundance errors in APOGEE DR17 (ses Table \ref{tab:uncertainties}). The black curves represent the ISM abundance tracks in the 8 kpc zone. The red contours represent a 2-D kernel density estimate of the APOGEE abundance distribution in that Galactic region, with the solid and dashed contours enclosing 30\% and 80\% of stars in the sample, respectively.}
    \label{fig:ofe-feh-sfh}
    \script{ofe_feh_sfh.py}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/ofe_feh_dtd.pdf}
    \caption{The [O/Fe]--[Fe/H] plane from multi-zone simulations with different DTD models. All assume the inside-out SFH. Each panel is similar to those in Figure \ref{fig:ofe-feh-sfh}, except each row contains star particles from a different bin in $|z|$, with stars closest to the midplane in the bottom row and stars farthest from the midplane in the top row. All panels present stars within the Solar annulus ($7\leq R_{\rm gal}<9$ kpc).}
    \label{fig:ofe-feh-dtd}
    \script{ofe_feh_dtd.py}
\end{figure*}

@@

We need to constrain DTD or galactic SFH better. With SDSS-V, the latter seems much more achievable.

@@

\begin{acknowledgments}
    Funding for the Sloan Digital Sky 
    Survey IV has been provided by the 
    Alfred P. Sloan Foundation, the U.S. 
    Department of Energy Office of 
    Science, and the Participating 
    Institutions.
